
# Data assembly {#data-assembly}

The following chapter details how we determined the area of interest and where we acquired the data products to be used in the connectivity analyses. We include justification of data sources where appropriate.  


```{r ch2_1, include=F}
# Global options

knitr::opts_chunk$set(
  comment = '', warning=F, message=F, echo=F)

```


```{r ch2_2, echo=F, include=F, warning=F}
#####################################################
###load packages###
# Install latest rgee
#remotes::install_github("r-spatial/rgee")
#remotes::install_github("r-earthengine/rgeeExtra")

list.of.packages <- c("stars",       
                      "sf",
                      "dplyr",            
                      "leaflet",
                      "rgee",
                      "rgeeExtra",
                      "elevatr",
                      "googledrive",
                      "purrr",
                      "remotes",
                      "reticulate",
                      "rnaturalearth",
                      "rmapshaper",
                      "leaflet",
                      "MetBrewer",
                      "ggmap",
                      "kableExtra",
                      "exactextractr",
                      "wdpar",
                      "terra")

# Check which ones you dont have
new.packages <- list.of.packages[!(list.of.packages %in% installed.packages()[,"Package"])]
#Install the ones you dont have
if(length(new.packages)) install.packages(new.packages)
# Load the packages
lapply(list.of.packages, require, character.only = TRUE)


# Authorise google drive
drive_auth()
1

```



```{r ch2_3, eval=F}

# Check rgee
# ONLY THE FIRST TIME: UNCOMMENT AND RUN THESE: Install the python packages:
#rgee::ee_install()
#rgee::ee_install_upgrade()

#py_install("numpy")


#rgee::ee_install_upgrade()
reticulate::py_install('earthengine-api==0.1.323')#, envname='PUT_HERE_YOUR_PYENV')

ee_Initialize()
# Check things are working
ee_check()
# Second text
reticulate::py_config()

```


```{r ch2_4, warning=F, message=F, include=F, eval=F}
# Download available shapefiles
dir.create("data")
dir.create("data/spatial")

#folder link to id
jp_folder = "https://drive.google.com/drive/folders/1YO0y73_jQ7Mo1fOnoI5EhCUDk5K9JEva"
folder_id = drive_get(as_id(jp_folder))

#find files in folder
files = drive_ls(folder_id)
i <- 1
#loop dirs and download files inside them
for (i in seq_along(files$name)) {
  #list files
  i_dir = drive_ls(files[i, ])
  
  #mkdir
  dir.create(paste0("data/spatial/",files$name[i]))
  
  #download files
  for (file_i in seq_along(i_dir$name)) {
    #fails if already exists
    try({
      drive_download(
        as_id(i_dir$id[file_i]),
        path = paste0("data/spatial/", files$name[i], "/", i_dir$name[file_i])
      )
    })
  }
}


# Download the low res version of Land cover map too

jp_folder = "https://drive.google.com/drive/folders/1YYaJ01gIMD9qs8cxKYqWmPk_lzyStDWy"
folder_id = drive_get(as_id(jp_folder))
files = drive_ls(folder_id)
for(i in 1:length(files))
{
drive_download(
        as_id(files$id[i]),
        path = paste0("data/spatial/landcover/", files$name[i]))
}

```

## Define the focal area

This project focuses on the area covered by the Mesoamerican Biological Corridor [MBC](https://en.wikipedia.org/wiki/Mesoamerican_Biological_Corridor), which spans most of mainland central America. This excludes islands in central America and the Caribbean, as these will likely need a differing set of ridge-to-reef definitions. We also exclude Mexico.   

```{r  ch2_5, echo=F, eval=F}
#devtools::install_github("ropensci/rnaturalearthhires")
library("rnaturalearthhires")

all_shp <- ne_countries(returnclass = "sf", scale=10)
all_shp <- all_shp[all_shp$subregion=="Central America",] 
# Remove mexico and clipperton
all_shp <- all_shp[all_shp$name_en!="Mexico",]
all_shp <- all_shp[all_shp$name_en!="Clipperton Island",]
all_shp<- st_transform(all_shp, 4326)
# Remove islands - but keep resolution
focal_shp <- ms_filter_islands(
  all_shp,
  min_area = 12391399903,
  drop_null_geometries = TRUE,
  force_FC = TRUE,
  sys_mem = 8
)

merged_shp <- st_union(focal_shp)

# Creater an exapnaded area for the "base aoi"

all_countries <- ne_countries(returnclass = "sf", scale=10)

# Use a 400km buffer for the base maps
buff_shp <- st_buffer(merged_shp, 400000)
#plot(buff_shp)
#plot(st_geometry(all_shp),add=T)
#plot(st_geometry(all_countries),add=T)

# Intersect islands with buffer
tmp <- st_intersection(all_countries, buff_shp)
# Remove islands
base_aoi <- ms_filter_islands(
  tmp,
  min_area = 12391399903,
  drop_null_geometries = TRUE,
  force_FC = TRUE,
  sys_mem = 8
)
base_aoi <- st_union(base_aoi)


#plot(st_geometry(base_aoi))
#plot(st_geometry(tmp))
#
# Write these files
st_write(all_shp,"data/spatial/area_of_interest/all_countries_islands.shp" )
st_write(focal_shp,"data/spatial/area_of_interest/focal_countries.shp" )
st_write(merged_shp, "data/spatial/area_of_interest/aoi.shp")
st_write(base_aoi, "data/spatial/area_of_interest/base_aoi.shp", append=F)


#plot(st_geometry(all_shp))
#plot(st_geometry(focal_shp))
#plot(st_geometry(merged_shp))


```


The full focal area spans:

```{r ch2_6}
all_shp <- st_read("data/spatial/area_of_interest/all_countries_islands.shp", quiet = TRUE)
focal_shp <- st_read("data/spatial/area_of_interest/focal_countries.shp", quiet = TRUE)
merged_shp <- st_read("data/spatial/area_of_interest/aoi.shp", quiet = TRUE)
base_shp <- st_read("data/spatial/area_of_interest/base_aoi.shp", quiet = TRUE)
base_terra <- vect("data/spatial/area_of_interest/base_aoi.shp")


leaflet(base_shp) %>%  
    addPolygons(
    stroke =T, color=met.brewer("Nizami")[4], 
    fillColor = met.brewer("Nizami")[4],
    fillOpacity = 0.5, smoothFactor = 0.5) %>%
  addPolygons(data=focal_shp, label = focal_shp$name,
    stroke =T, color=met.brewer("Nizami")[2], 
    fillColor = met.brewer("Nizami")[2],
    fillOpacity = 0.5, smoothFactor = 0.5) %>% 
  addProviderTiles("Esri.WorldPhysical") %>% 
  addLegend(position="topright", labels=c("focal area", "simulated area"), colors=c(met.brewer("Nizami")[2], met.brewer("Nizami")[4]), title="Area of interest")

# Setup the boundry for GEE
mbc_bbox <- st_bbox(base_shp) # changed from focal_shp

(mbc_bbox[2]+mbc_bbox[4])/2
#12.6 - have to go lower

boundary_upper <- mbc_bbox
boundary_lower <- mbc_bbox

boundary_upper[2] <- 8
boundary_lower[4] <- 8

```

The area in orange is the region which we aquire data for, the red area represents the regions for which we explore the connectivity. The reason for the expansion is to minimise the possibility of edge effects. 


```{r  ch2_7, eval=F}
# Subset using the area of interest
focal <- ee$Geometry$Rectangle(as.numeric(st_bbox(mbc_bbox)))

focal_upper <- ee$Geometry$Rectangle(as.numeric(st_bbox(boundary_upper)))
focal_lower <- ee$Geometry$Rectangle(as.numeric(st_bbox(boundary_lower)))



#Map.addLayer(focal)
```


Below we show the exclusion of surrounding islands:

```{r ch2_8}
# Island exclusions
plot(st_geometry(all_shp), col=met.brewer("Nizami")[3], border=F)
plot(st_geometry(focal_shp), col=met.brewer("Nizami")[5], add=T, border=F)
legend("topleft", c("Included", "Excluded"), col=c(met.brewer("Nizami")[5], met.brewer("Nizami")[3]), pch=15, cex=1)

```


```{r ch2_9, eval=F, include=F}
# Check this also overlaps with the MBC
CBM_shp<- st_read("data/spatial/CBM Regional/cbm_regional.shp", quiet = TRUE)
plot(st_geometry(merged_shp))
plot(st_geometry(CBM_shp), add=T, col="red")
#str(CBM_shp)
#plot(st_geometry(CBM_shp), col="red")

```


For an up to date assessment of the Mesoamerican Biological Corridor in Panama alone see:

[Meyer, N. F., Moreno, R., Reyna-Hurtado, R., Signer, J., & Balkenhol, N. (2020). Towards the restoration of the Mesoamerican Biological Corridor for large mammals in Panama: comparing multi-species occupancy to movement models. Movement ecology, 8(1), 1-14.](https://movementecologyjournal.biomedcentral.com/articles/10.1186/s40462-019-0186-0)


## Data products

### Protected areas
Shape files for protected areas were downloaded from the Protected Planet database. The World Database on Protected Areas (WDPA) is the most up-to-date and complete source of information on protected areas, updated monthly with submissions from governments, non-governmental organizations, landowners, and communities. It is managed by the United Nations Environment Programme's World Conservation Monitoring Centre (UNEP-WCMC) with support from IUCN and its World Commission on Protected Areas (WCPA).

We buffered the area of interest by 10km, then excluded any protected ares which fell outside of that zone. This means marine protected areas >10km from the shore are not considered. There are two broad types of park - National and International designations - there are also many further subdivisions not considered here.

As recommended in the WDPA best practices guide, we removed any PA that did not report its area, or with a ‘Proposed’ status or ‘UNESCO-MAB Biosphere Reserve’ designation (Note core areas remain under national park designations).

We now also use the package `wdpar` to clean known issues in the database (i.e. filter out overlapping sections).

```{r  ch2_10, echo=F, eval=F}
gpa  <- ee$FeatureCollection("WCMC/WDPA/current/polygons")

gpa_upper <- gpa$filterBounds(focal_upper)
gpa_lower <- gpa$filterBounds(focal_lower)

#Veiw on google earth
#Map$setCenter(-85, 10, 4)
#Map$addLayer(gpa_upper)
Map$addLayer(gpa)

# My layer is too large to convert to sf without filtering
# Filter to get only larger continental US watersheds.
#gpa_filtered$filter(ee$gpa_filtered$gt("areasqkm", 25000))

gpa_upper_shp <- ee_as_sf(gpa_upper)    
gpa_lower_shp <- ee_as_sf(gpa_lower)    

gpa_shp <- rbind(gpa_upper_shp, gpa_lower_shp)

library(wdpar)

gpa_clean <- wdpa_clean(
  gpa_shp,
  crs = paste("+proj=cea +lon_0=0 +lat_ts=30 +x_0=0",
    "+y_0=0 +datum=WGS84 +ellps=WGS84 +units=m +no_defs"),
  exclude_unesco = TRUE,
  retain_status = c("Designated", "Inscribed", "Established"),
  snap_tolerance = 1,
  simplify_tolerance = 0,
  geometry_precision = 1500,
  erase_overlaps = TRUE,
  verbose = interactive()
)

gpa_clean <- st_transform(gpa_clean, 4326)


leaflet() %>%
  addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
  addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
  addPolygons(data=gpa_clean, fillColor = "blue",
               label = gpa_clean$NAME, group="Protected areas", stroke=T) %>%
  addLayersControl(
    baseGroups = c("OSM (default)", "Satellite"),
    overlayGroups = c("Protected areas"))


# 
# # Remove duplicated geometries
# gpa_clean <- gpa_clean[duplicated(gpa_clean$NAME)==F,]
# # Remove the regional designations
# gpa_clean <- gpa_clean[gpa_clean$DESIG!="Regional",]
# 
# # Remove those that are not terrestrial - 10km of land
# gpa_ter <- gpa_clean[(gpa_clean$GIS_AREA-gpa_clean$GIS_M_AREA)>0,]
# 
# # Remove parks without areas
# gpa_ter <- gpa_ter[gpa_ter$GIS_AREA>0.01,]
# 
# # Calculate terrestrial area
# gpa_ter$GIS_T_AREA <- gpa_ter$GIS_AREA -  gpa_ter$GIS_M_AREA
# 
# summary(gpa_ter$GIS_AREA)
# summary(gpa_ter$GIS_T_AREA)
# 
# # Remove Proposed status
# gpa_ter <- gpa_ter[gpa_ter$STATUS!="Proposed",]
# 
# # Remove UNESCO-MAB Biosphere Reserve
# gpa_filtered <- gpa_ter[gpa_ter$DESIG!="UNESCO-MAB Biosphere Reserve",]
# plot(st_geometry(gpa_filtered))
# 
# # Remove known duplicated
# # Rio Platano
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Laguna de Términos",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Patanos de Centia",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Darien National Park",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Los Katios",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Talamanca Range-La Amistad Reserves / La Amistad National Park",]
# 
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Turberas de Talamanca",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Sistema de Humedales de la Bahía de Bluefields",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Complejo Bahía de Jiquilisco",]
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Río Plátano Biosphere Reserve",]
# 
# gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Corredor Fronterizo",]
# 
# 
# # Temp
# #gpa_filtered <- gpa_filtered[gpa_filtered$NAME != "Tapanti-Macizo de la Muerte",]
# #
# 
# 	
# 
# # temp 	
# # New idea, order by name and size, if duplicated take the first
# gpa_filtered <- gpa_filtered[order(gpa_filtered$NAME, gpa_filtered$GIS_AREA, decreasing = T),]
# 
# #gpa_filtered[gpa_filtered$NAME=="Golfo de Montijo",]
# gpa_filtered <- gpa_filtered[duplicated(gpa_filtered$NAME)==F,]
# 
# 
# leaflet() %>%
#   addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
#   addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
#   addPolygons(data=gpa_filtered[st_geometry_type(gpa_filtered)!="GEOMETRYCOLLECTION",], fillColor = "blue",
#                label = gpa_filtered$NAME[st_geometry_type(gpa_filtered)!="GEOMETRYCOLLECTION"], group="Protected areas", stroke=T) %>%
#   addLayersControl(
#     baseGroups = c("OSM (default)", "Satellite"),
#     overlayGroups = c("Protected areas"))
# 
# leaflet() %>%
#   addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
#   addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
#   addPolygons(data=gpa_clean, fillColor = "blue",
#                label = gpa_clean$NAME, group="Protected areas", stroke=T) %>%
#   addLayersControl(
#     baseGroups = c("OSM (default)", "Satellite"),
#     overlayGroups = c("Protected areas"))
# 
# 
# # CAN I MEGRE OVERLAPPING POLYGONS
# 
# st_overlaps(gpa_filtered, sparse=F)
# ?st_overlaps

# Subset to focal protected areas

tmp <- st_transform(merged_shp,3160)
tmp2 <- st_buffer(tmp,50)
#plot(st_geometry(tmp))
#plot(st_geometry(tmp2))
tmp2 <- st_transform(tmp2, 4326)

gpa_clean <- st_make_valid(gpa_clean)
sf_use_s2(FALSE)
gpa_clean <- sf::st_buffer(gpa_clean, dist = 0)


# Interset with the core area
#focal_pa <- st_intersection(gpa_clean, tmp2)
focal_pa <- st_intersection(gpa_clean, merged_shp)
plot(st_geometry(focal_pa))

leaflet() %>%
  addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
  addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
  addPolygons(data=focal_pa, fillColor = "blue",
               label = focal_pa$NAME, group="Protected areas", stroke=T) %>%
  addLayersControl(
    baseGroups = c("OSM (default)", "Satellite"),
    overlayGroups = c("Protected areas"))

?st_area
#hist(st_area(focal_pa)/(1000000))
table(as.numeric(st_area(focal_pa)/(1000000))<0.5)

focal_pa <- focal_pa[as.numeric(st_area(focal_pa)/(1000000))>0.5,]

# Remove the weird coastal small shit

focal_pa <- focal_pa[!(focal_pa$NAME %in% c( "Complejo Los Cóbanos", "Playa Boca Vieja", 	
"Manglares de Panamá Viejo", "Nacional Cariari", "Cayos Perlas", "Port Honduras", "Corozal Bay")),]

# There are two Alto darien's - remove other duplicates
focal_pa <- focal_pa[duplicated(focal_pa)==F,]
focal_pa <- focal_pa[duplicated(focal_pa$geometry)==F,]

# leaflet() %>%
#   addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
#   addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
#   addPolygons(data=focal_pa, fillColor = "blue",
#                label = focal_test$NAME, group="Protected areas", stroke=T) %>%
#   addLayersControl(
#     baseGroups = c("OSM (default)", "Satellite"),
#     overlayGroups = c("Protected areas"))


#test <- st_combine(focal_pa, by_feature=T)
# ?st_combine
# ?st_cast
# st_cast(focal_pa, "P")
# # Union the shapes
# focal_merged <- st_union(focal_pa, by_feature=T)
# focal_pa
# head(focal_merged)
# 
# plot(test)
# leaflet(test)
# 
# g <- st_intersects(focal_pa,focal_pa)



# Store this file
dir.create("data/spatial/WDPA_protected_areas")
saveRDS(focal_pa, "data/spatial/WDPA_protected_areas/focal_area_pa_shp.RDS")

drive_upload("data/spatial/WDPA_protected_areas/focal_area_pa_shp.RDS", as_id("https://drive.google.com/drive/u/0/folders/1KgNuMrKPyskUmroOBGISt3FxAsaBLmRv"))

```

```{r  ch2_11}
# Plot
focal_pa <- readRDS( "data/spatial/WDPA_protected_areas/focal_area_pa_shp.RDS")
plot(st_geometry(merged_shp), col=rgb(0,0,0,0.1), border=F)
plot(st_geometry(focal_pa[focal_pa$DESIG_TYPE=="International",]), col=met.brewer("Nizami")[1], add=T)
plot(st_geometry(focal_pa[focal_pa$DESIG_TYPE=="National",]), col=met.brewer("Nizami")[6], add=T)
legend("topright", c("National", "International"), pch=15, col=c(met.brewer("Nizami")[6], met.brewer("Nizami")[1]))

```

Data source:  UNEP-WCMC and IUCN (year), Protected Planet: The World Database on Protected Areas (WDPA) [February 2022], Cambridge, UK: UNEP-WCMC and IUCN Available at: [Protected Planet](www.protectedplanet.net).

And and all protected areas in an interactive version:

```{r ch2_11b}
#head(focal_pa)
#table(st_geometry_type(focal_pa))

leaflet() %>%
  addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
  addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
  addPolygons(data=focal_pa[st_geometry_type(focal_pa)!="GEOMETRYCOLLECTION",], fillColor = "blue",
               label = focal_pa$NAME[st_geometry_type(focal_pa)!="GEOMETRYCOLLECTION"], group="Protected areas", stroke=F) %>%
  addLayersControl(
    baseGroups = c("OSM (default)", "Satellite"),
    overlayGroups = c("Protected areas"))
```

#### Types of protected area
Within our focal area, the WPDA dataset includes the following number of terrestrial and marine protected areas:

```{r  ch2_12, echo=F}
tmp <- as.data.frame(table(focal_pa$MARINE))
tmp  %>%
  kbl() %>%
  kable_styling()
```

There are also a myriad of different protection designations:

```{r  ch2_13, echo=F}

tmp <- data.frame(table(focal_pa$DESIG))
colnames(tmp) <- c("Type", "Freq")
tmp  %>%
  kbl() %>%
  kable_styling() %>%
  scroll_box(width = "100%", height = "400px")

```

### Elevation
We downloaded the elevation of the area of interest using SRTM Digital Elevation Data Version 4. The Shuttle Radar Topography Mission (SRTM) digital elevation dataset was originally produced to provide consistent, high-quality elevation data at near global scope. 

```{r  ch2_15, eval=F, include=F}
# First time only
dem <- ee$Image("CGIAR/SRTM90_V4")
#Map$setCenter(-122.1899, 37.5010, 10)
#Map$addLayer(dem)
#focal

#      xmin       ymin       xmax       ymax 
#-96.067919   3.645258 -73.465162  21.623440 

(3.645258 + 21.623440)/2

# Upper
geom_params <-   list(
  scale = 400,
  crs = 'EPSG:4326',
  region = '[[-96.067919 , 3.645258], [-73.465162, 3.645258], [-73.465162, 12.6], [-96.067919 , 12.6]]'
)
path <- dem$getDownloadUrl(geom_params)

# SAVE THE FILES TO YOUR GOOGLE DRIVE MANUALLY (DONE)

# Lower
geom_params <-   list(
  scale = 400,
  crs = 'EPSG:4326',
  region = '[[-96.067919 , 12.6], [-73.465162, 12.6], [-73.465162, 21.623440], [-96.067919 , 21.623440]]'
)
path <- dem$getDownloadUrl(geom_params)



remove(dem)
```


```{r ch2_16}
# First time run

# Read in upper
#dem_raster1 <- read_stars("data/spatial/SRTM90_V4/SRTM90_V4.elevation_upper.tif")
# Read in lower
#dem_raster2 <- read_stars("data/spatial/SRTM90_V4/SRTM90_V4.elevation_lower.tif")

# Stitch together
#dem_raster <- st_mosaic(dem_raster1, dem_raster2)

# dem_raster <- read_stars("data/spatial/SRTM90_V4/SRTM90_V4.elevation.tif")
# # Mask out the areas you are not interested in
#dem_raster <- st_crop(dem_raster, base_shp)
# plot(dem_raster)

# write_stars(dem_raster, "data/spatial/SRTM90_V4/SRTM90_V4.elevation_all.tif")
library(terra)
dem_raster <- rast("data/spatial/SRTM90_V4/SRTM90_V4.elevation_all.tif")
dem_crop <- mask(dem_raster, base_terra)
writeRaster(dem_crop, "data/spatial/SRTM90_V4/SRTM90_V4.elevation_mask.tif")
plot(dem_crop)

# # Mask out the areas you are not interested in
# gfc_2020 <- st_crop(gfc_2020, base_shp)
# write_stars(gfc_2020, "data/spatial/gfc/global_forest_change_2020_v1_8.treecover2000.tif")

#dem_raster <- read_stars("data/spatial/SRTM90_V4/SRTM90_V4.elevation.tif")
#plot(dem_raster, las=1,nbreaks=11 ,breaks="equal", col=terrain.colors(10), bg="grey", main="")
```

Data source: [Jarvis, A., H.I. Reuter, A. Nelson, E. Guevara. 2008. Hole-filled SRTM for the globe Version 4, available from the CGIAR-CSI SRTM 90m Database](https://srtm.csi.cgiar.org).



###	Forest cover (current)
To get a layer reflecting current forest cover we use the Hansen Global forest Change index. These reflect results from time-series analysis of Landsat images in characterizing global forest extent and change. This data is up to date until 2021!


```{r  ch2_17, include=F, eval=F}
# First run only 

gfc <- ee$Image("UMD/hansen/global_forest_change_2021_v1_9")
Map$addLayer(
    eeObject = gfc,
    visParams = list(
      bands = c('treecover2000'),
      min = 0,
      max = 100,
      palette = c('white','green')),
    name = 'Forest cover')


# Map$addLayer(
#     eeObject = gfc,
#     visParams = list(
#       bands = c('loss'),
#       min = 0,
#       max = 1,
#       palette = c('white','red')),
#     name = 'Forest cover')


loss <- gfc$select("loss")
notLoss <- (loss[["loss"]] *(-1)) +1

gfc_remain <- gfc$updateMask(notLoss)

Map$addLayer(
    eeObject = gfc_remain,
    visParams = list(
      bands = c('treecover2000'),
      min = 0,
      max = 100,
      palette = c('white','green')),
    name = 'Forest cover')

#      xmin       ymin       xmax       ymax 
#-96.067919   3.645258 -73.465162  21.623440 

# Lower
geom_params <-   list(
  scale = 400,
  crs = 'EPSG:4326',
  region = '[[-96.067919 , 3.645258], [-73.465162, 3.645258], [-73.465162, 12.6], [-96.067919 , 12.6]]'
)
path <- gfc_remain[["treecover2000"]]$getDownloadUrl(geom_params)

# SAVE THE FILES TO YOUR GOOGLE DRIVE MANUALLY (DONE)

# Upper
geom_params <-   list(
  scale = 400,
  crs = 'EPSG:4326',
  region = '[[-96.067919 , 12.6], [-73.465162, 12.6], [-73.465162, 21.623440], [-96.067919 , 21.623440]]'
)
path <- gfc_remain[["treecover2000"]]$getDownloadUrl(geom_params)

# PASTE THAT INTO YOUR BROWSER

# IF YOU NEED HIGH RESOILUTION DO IT LIKE THIS

#ee_as_raster(gfc_remain[["treecover2000"]], maxPixels = 10000000000, scale=400)

```

```{r  ch2_18}
# Run hashed files the first time only
# gfc_2020_upper <- rast("data/spatial/gfc/global_forest_change_2021_v1_9.treecover2000_upper.tif")
# gfc_2020_lower <- rast("data/spatial/gfc/global_forest_change_2021_v1_9.treecover2000_lower.tif")
# m2 <- mosaic(gfc_2020_upper, gfc_2020_lower)
# gfc2020 <- mask(m2, base_terra)
# plot(gfc2020)
# writeRaster(gfc2020, "data/spatial/gfc/global_forest_change_2021_v1_9.treecover2000.tif")

gfc_2020 <- rast("data/spatial/gfc/global_forest_change_2021_v1_9.treecover2000.tif")
plot(gfc_2020)

```

Possible future data incorporation: 

Current (and future) coarse vegetation types can also be obtained from this [Baumbach, L., Warren, D. L., Yousefpour, R., & Hanewinkel, M. (2021). Climate change may induce connectivity loss and mountaintop extinction in Central American forests. Communications Biology, 4(1), 1-12.](https://www.nature.com/articles/s42003-021-02359-9) 

For an example approaches in modelling connectivity in the future (beyond the remit of this contract) see: [Mozelewski, T. G., Robbins, Z. J., Scheller, R. M., & Mozelewski, T. G. (2022). Forecasting the influence of conservation strategies on landscape connectivity. Conservation Biology](https://conbio.onlinelibrary.wiley.com/doi/epdf/10.1111/cobi.13904)


###	Forest biomass
We obtained above ground biomass from [Spawn, S.A., Sullivan, C.C., Lark, T.J. et al. Harmonized global maps of above and belowground biomass carbon density in the year 2010. Sci Data 7, 112 (2020)](https://doi.org/10.1038/s41597-020-0444-4). This dataset provides temporally consistent and harmonized global maps of above-ground and below-ground biomass carbon density for the year 2010 at a 300-m spatial resolution.

The values represent above-ground living biomass carbon stock density of combined woody and herbaceous cover in 2010. This includes carbon stored in living plant tissues that are located above the earth’s surface (stems, bark, branches, twigs). This does not include leaf litter or coarse woody debris that were once attached to living plants but have since been deposited and are no longer living.

```{r  ch2_19, include=F, eval=F}
# First run only 
biomass <- ee$ImageCollection("NASA/ORNL/biomass_carbon_density/v1")
test <- biomass$select("agb")
test <- ee_imagecollection_to_local(
  test,
  region = focal,
  dsn = NULL,
  via = "drive",
  container = "rgee_backup",
  scale = NULL,
  maxPixels = 1e+09
)
drive_auth()



?ee_imagecollection_to_local
drive_auth(email = "c.w.beirne@gmail.com")
#drive_auth()
nasa_biomass <- read_stars("data/spatial/biomass/NASA_biomass_desnity_estimation.tif")
# Mask out the areas you are not interested in
nasa_biomass <- st_crop(nasa_biomass, merged_shp)
write_stars(nasa_biomass, "data/spatial/biomass/NASA_biomass_desnity_estimation.tif")

```

```{r ch2_20}
nasa_biomass <- read_stars("data/spatial/biomass/NASA_biomass_desnity_estimation.tif")
plot(nasa_biomass, las=1,nbreaks=11 ,breaks="equal", col=rev(terrain.colors(10)), bg="grey", main="")
```

### Forest height
For forest height we use the global 2005 dataset  representing global tree heights based on a fusion of spaceborne-lidar data (2005) from the Geoscience Laser Altimeter System (GLAS) and ancillary geospatial data. See Simard et al. (2011) for details. 

[Simard, M., Pinto, N., Fisher, J., Baccini, A. 2011. Mapping forest canopy height globally with spaceborne lidar. Journal of Geophysical Research. 116: G04021](https://agupubs.onlinelibrary.wiley.com/doi/10.1029/2011JG001708)

```{r  ch2_21, include=F, eval=F}
height <- ee$Image("NASA/JPL/global_forest_canopy_height_2005")
geom_params <-   list(
  scale = 400,
  crs = 'EPSG:4326',
  region = '[[-92.246257 , 7.205715], [-77.163270, 7.205715], [-77.163270, 18.490759], [-92.246257 , 18.490759]]'
)
path <- height$getDownloadUrl(geom_params)
# sAVE THE FILE TO YOUR GOOGLE DRIVE MANUALLY (DONE)
remove(dem)

height <- read_stars("data/spatial/height/global_forest_canopy_height_2005.1.tif")
# Mask out the areas you are not interested in
height <- st_crop(height, merged_shp)
write_stars(height, "data/spatial/height/global_forest_canopy_height_2005.1.tif")
```


```{r ch2_22}
height <- read_stars("data/spatial/height/global_forest_canopy_height_2005.1.tif")
plot(height, las=1,nbreaks=11 ,breaks="equal", col=rev(terrain.colors(10)), bg="grey", main="")

```

### Mangrove cover (current)

Mangrove data is taken from the [USGS: Global Distribution of Mangroves]  (https://data.unep-wcmc.org/datasets/4). 

```{r  ch2_23, eval=F}
# Original run -> file too big to keep
man_shp <- st_read("data/spatial/WCMC010_MangrovesUSGS2011_v1_4/14_001_WCMC010_MangroveUSGS2011_v1_4.shp", quiet = TRUE)
man_shp <- st_transform(man_shp,4326)
tmp <- st_transform(merged_shp,3160)
tmp2 <- st_buffer(tmp,10000)
plot(st_geometry(tmp))
plot(st_geometry(tmp2))
tmp2 <- st_transform(tmp2, 4326)
focal_mangrove <- st_intersection(man_shp, tmp2)
st_write(focal_mangrove, "data/spatial/WCMC010_MangrovesUSGS2011_v1_4/clipped/mangroves_clipped.shp", quiet = TRUE)
remove(man_shp)
```


```{r ch2_24}
focal_mangrove <- st_read("data/spatial/WCMC010_MangrovesUSGS2011_v1_4/clipped/mangroves_clipped.shp", quiet = TRUE)
plot(st_geometry(merged_shp), col=rgb(0,0,0,0.1), border=F)
plot(st_geometry(focal_mangrove), add=T, col=met.brewer("Nizami")[2], border=F)
legend("topright",c("Mangrove","Land"), pch=15, col=c(met.brewer("Nizami")[2],rgb(0,0,0,0.1)))

```

Citation:
Giri C, Ochieng E, Tieszen LL, Zhu Z, Singh A, Loveland T, Masek J, Duke N (2011). Status and distribution of mangrove forests of the world using earth observation satellite data (version 1.4, updated by UNEP-WCMC). Global Ecology and Biogeography 20: 154-159. Paper DOI: 10.1111/j.1466-8238.2010.00584.x . Data DOI: https://doi.org/10.34892/1411-w728


### Human disturbance:  

To incorporate human disturbance into the connectivity analyses we use the global Human Modification dataset (gHM). This dataset provides a cumulative measure of human modification of terrestrial lands globally at 1 square-kilometer resolution. The gHM values range from 0.0-1.0 and are calculated by estimating the proportion of a given location (pixel) that is modified, the estimated intensity of modification associated with a given type of human modification or "stressor". They mapped 5 major anthropogenic stressors circa 2016 were mapped using 13 individual datasets:

- human settlement (population density, built-up areas)
- agriculture (cropland, livestock)
- transportation (major, minor, and two-track roads; railroads)
- mining and energy production
- electrical infrastructure (power lines, nighttime lights)

As such, this layer represents a great starting point to measure broad scale patterns in elevational gradient disturbances. 

[Kennedy, C.M., J.R. Oakleaf, D.M. Theobald, S. Baurch-Murdo, and J. Kiesecker. 2019. Managing the middle: A shift in conservation priorities based on the global human modification gradient. Global Change Biology 00:1-16.](https://onlinelibrary.wiley.com/doi/10.1111/gcb.14549)

```{r  ch2_25, eval=F}
# First time only
hmi <- ee$ImageCollection("CSP/HM/GlobalHumanModification")
test <- hmi$select("gHM")
test <- ee_imagecollection_to_local(
  test,
  region = focal,
  dsn = NULL,
  via = "drive",
  container = "rgee_backup",
  scale = NULL,
  maxPixels = 1e+09
)
```


```{r ch2_26}
# First time run
# ghm_raster <- read_stars("data/spatial/human_modification_index/gHM_2016.tif")
# # Mask out the areas you are not interested in
# ghm_raster <- st_crop(ghm_raster, merged_shp)
# write_stars(ghm_raster, "data/spatial/human_modification_index/gHM_2016.tif")

ghm_raster <- read_stars("data/spatial/human_modification_index/gHM_2016.tif")

plot(ghm_raster, las=1,nbreaks=11 ,breaks="equal", col=terrain.colors(10), bg="grey")

```


Examples of papers usuing the human modification index (or a derivation of it):
- Gray M, Micheli E, Comendant T, Merenlender A (2020) Quantifying climate-wise connectivity across a topographically diverse landscape. Land 9:1–18. https://doi.org/10.3390/land9100355
      This paper does terrestrial and riparian "permeability" - the inverse of resistance. They then compare the "cooling potential" of these corridors through looking at the pairwise difference between temperatures between linkages. 



### Current land-use and habitat

To capture land use, we are used the high resolution (10m) land cover map over Mexico and Central America created by ESA. The data are based on more than 2 years of Sentinel-2A and 2B observations from January 2016 to March 2018.


```{r  ch2_27, echo=F}
lc_raster <- read_stars("data/spatial/landcover/ESACCI-LC-L4-LC10-Map-10m-CAM.tif")
lc_key <- read.csv("data/spatial/landcover/ESACCI-LC_S2_Prototype_ColorLegend.csv", header=T, sep=";")
lc_key$rgb <- rgb(lc_key$R/256,lc_key$G/256,lc_key$B/256)
# Remove no data
lc_key <- lc_key[1:11,]
# Simplify labels
lc_key$LCCOwnLabel[lc_key$LCCOwnLabel=="Tree cover areas"] <- "Tree cover"
lc_key$LCCOwnLabel[lc_key$LCCOwnLabel=="Shrubs cover areas"] <- "Shrub cover"
lc_key$LCCOwnLabel[lc_key$LCCOwnLabel=="Vegetation aquatic or regularly flooded"] <- "Flooded vegetation"

# Mask out the areas not in the focal region
#plot(st_geometry(merged_shp))

lc_raster <- st_crop(lc_raster, merged_shp, crop=FALSE)

plot(lc_raster, breaks=seq(-0.5,10.5, by=1),col=lc_key$rgb[1:11], key.pos=NULL)
legend("topright",legend=lc_key$LCCOwnLabel, fill=lc_key$rgb )
```

## Start and end nodes

The analysis presented in the following chapters depends on having a suite of meaning start locations "reef" (reflecting mangroves, coastal protected areas and marine protected areas) and end locations "ridges" (reflecting protected high elevation forest habitats). The following code outlines the candidate start and end locations. 

### Start nodes

This is where the connectivity paths will start from:

#### Mangroves >1km2

According to the USGS: Global Distribution of Mangroves dataset there are `r nrow(focal_mangrove)` discrete mangrove patches across the study area. The vast majority of these, however, are small:

```{r ch2_28}
hist(focal_mangrove$AREA_KM2, las=1, ylab="", xlab="Size of fragment (km2)", main="")    
```

The distribution of sizes (km2) is as follows: 

```{r ch2_29}
summary(focal_mangrove$AREA_KM2)
```

Consequently, we only focus on large mangrove fragments (e.g. > 1 square kilometers) as this represents the resolution of our input layer. We have merged fragments which are very close to one another (e.g. on opposite sides of the river).

```{r  ch2_30, eval=F, include=F}
# Buffer mangroves by 500m, then merge them to see how it influences size distribution. 
focal_mangrove_utm <- st_transform(focal_mangrove, 3160)

# Buffer by 100m (toremove fragments)
focal_mangrove_utm_buff <- st_buffer(focal_mangrove_utm, 100)

# Merge them
focal_mangrove_utm <- st_union(focal_mangrove_utm_buff)

# Reconvert back to individual polygons
focal_mangrove_utm <- st_cast(focal_mangrove_utm,"POLYGON")

# Then back to simple features
focal_mangrove_utm <- st_as_sf(focal_mangrove_utm)

# Add the areas
focal_mangrove_utm$AREA_M2 <- as.numeric(st_area(focal_mangrove_utm))
focal_mangrove_utm$AREA_KM2 <- as.numeric(st_area(focal_mangrove_utm)/1000000)

# Convert back to lat/long
focal_mangrove <- st_transform(focal_mangrove_utm, 4326)

# Subset to patches > 1km2
mangrove_large <- focal_mangrove[focal_mangrove$AREA_KM2>1,]

#leaflet() %>% 
#  addTiles() %>% 
#  addPolygons(data=mangrove_large, weight = 2, fillColor = "blue")
# Write the shapefile

# Add back in the mangrove names (if they exist)

st_write(mangrove_large, "data/spatial/mangroves/mangrove_large_1kmsq.shp", overwrite=T, append=F)


```

```{r  ch2_31,include=F, message=F, warning=F}
mangrove_large <- st_read("data/spatial/mangroves/mangrove_large_1kmsq.shp")
```

If we subset to just fragments greater than 1 km2, we have `r nrow(mangrove_large)` fragments. The centroids of these "large mangrove" patches are distributed as follows:

```{r  ch2_32, echo=F}

#leaflet() %>% 
#  addTiles() %>% 
#  addPolygons(data=tmp, weight = 2, fillColor = "red") %>% 
#  addPolygons(data=mangrove_large, weight = 2, fillColor = "blue")

plot(st_geometry(merged_shp), col=rgb(0,0,0,0.1), border=F)
#plot(mangrove_large, add=T, col=met.brewer("Nizami")[2], border=F)
plot(st_geometry(mangrove_large), pch=19, col=paste0(met.brewer("Nizami")[2], "50"), add=T, border=met.brewer("Nizami")[2])
legend("topright",c("Mangrove","Land"), pch=15, col=c(met.brewer("Nizami")[2],rgb(0,0,0,0.1)))

```

#### Coastal Protected areas

The distribution of coastal protected areas which overlap (i.e. are within 1km) the coast is as follows:


```{r,  ch2_33, echo=F, eval=F, message=F}
sf_use_s2(FALSE)
# First donwload a line feature that encompasses the coast
#plot(st_geometry(merged_shp))

coast_shp <- ne_coastline(scale = 10, returnclass = c( "sf"))
#plot(st_geometry(coast_shp), add=T, col="blue")

# Buffer the focal area they crop the coastline layer (to remove islands etc)
tmp <- st_transform(merged_shp, 3160 )
tmp <- st_buffer(tmp, 500)
tmp <- st_transform(tmp, 4326)

focal_coast <- st_intersection(coast_shp, tmp)
#plot(st_geometry(focal_coast), add=T, col="red", lwd=2)

# Buffer pa'S by 1000m then intersect with protected areas
tmp <- st_transform(focal_pa, 3160 )
tmp <- st_buffer(tmp, 1000)
tmp <- st_transform(tmp, 4326)
#plot(st_geometry(tmp), add=T)

# Tried to intersect with the focal coastline but it didnt work 
test <- focal_pa[st_intersects(tmp, focal_coast, sparse=F),]
plot(st_geometry(merged_shp))

# This creates polygons of the protected area strips
test <- st_intersection(tmp, focal_coast)
#plot(st_geometry(merged_shp))
#plot(st_geometry(test), add=T, col="green")


coastal_pa <- focal_pa[unlist(st_intersects(test,focal_pa, sparse=T)),]
#plot(st_geometry(coastal_pa), add=T, col="green")

# Remove duplicates
coastal_pa <- coastal_pa[duplicated(coastal_pa)==F,]
plot(st_geometry(coastal_pa), add=T, col="red")

leaflet() %>%
  addProviderTiles("OpenStreetMap", group="OSM (default)") %>%
  addProviderTiles("Esri.WorldImagery", group="Satellite") %>%
  addPolygons(data=coastal_pa[st_geometry_type(coastal_pa)!="GEOMETRYCOLLECTION",], fillColor = "blue",
               label = coastal_pa$NAME[st_geometry_type(coastal_pa)!="GEOMETRYCOLLECTION"], group="Protected areas") %>%
  addLayersControl(
    baseGroups = c("OSM (default)", "Satellite"),
    overlayGroups = c("Protected areas"))




# Remove parks without areas
coastal_pa <- coastal_pa[coastal_pa$GIS_AREA>0.01,]
coastal_pa <- coastal_pa[coastal_pa$REP_AREA>0.01,]

# Remove parks < 1km2
coastal_pa <- coastal_pa[coastal_pa$GIS_AREA>1,]

# Remove Proposed status
coastal_pa <- coastal_pa[coastal_pa$STATUS!="Proposed",]
# Remove UNESCO-MAB Biosphere Reserve
coastal_pa <- coastal_pa[coastal_pa$DESIG!="UNESCO-MAB Biosphere Reserve",]

plot(st_geometry(coastal_pa[st_geometry_type(coastal_pa)=="GEOMETRYCOLLECTION",]), add=T, col="blue")

coastal_pa <- coastal_pa[st_geometry_type(coastal_pa)!="GEOMETRYCOLLECTION",]

#Remove the border park
coastal_pa <- coastal_pa[coastal_pa$NAME!="Corredor Fronterizo",]

st_write(coastal_pa, "data/spatial/protected_areas/Coastal_protected_areas.shp", overwrite=T, append=F)

```


```{r, echo=F}
coastal_pa <- st_read("data/spatial/protected_areas/Coastal_protected_areas.shp")

plot(st_geometry(merged_shp))
plot(st_geometry(coastal_pa),add=T, col="red")
```

So we have around `r nrow(coastal_pa)` protected areas which are link ocean and land. Note this also includes near-shore marine PA's. For each of these locations, it doesnt make sense to include the full protected area within these calculations as they sometimes run a long way inland (biasing connectivity estimates). The center points of the protected areas would not reflect reef or ocean connectivity. We therefore crop each of these protected areas to the zones within 1 km of the coast, and will use these as connectivity start points.

Map of the protected areas with 1km of the coast:

```{r  ch2_34, eval=F, include=F, message=F, warning=F}
sf_use_s2(FALSE)
# First buffer to terrestrial <500m from the ocean
# Buffer pa'S by 1000m then intersect with protected areas
tmp <- st_transform(focal_coast, 3160)
tmp <- st_buffer(tmp, 1000)
tmp <- st_transform(tmp, 4326)

# Intersect the coasta PAS' with buffer
test <- st_intersection(coastal_pa, tmp )

# NOTE THIS GIVES YOU EVERYTHING, we need to mask out the water areas
test2 <- st_intersection(test, focal_shp )


# Add it to a leaflet to check
leaflet() %>% 
  addTiles() %>% 
  addPolygons(data=test, weight = 2, fillColor = "red") %>% 
  addPolygons(data=test2, weight = 2, fillColor = "blue")

# If we have multiple fragments for each prtected area - take the largest
test2$COAST_AREA <- st_area(test2)

test2 <- test2[order(test2$NAME, test2$COAST_AREA),]
test3 <- test2[duplicated(test2$NAME)==F,]


# leaflet() %>% 
#   addTiles() %>% 
#   addPolygons(data=test, weight = 2, fillColor = "red") %>% 
#   #addPolygons(data=test2, weight = 2, fillColor = "blue") %>% 
#   addPolygons(data=test3, weight = 5, fillColor = "black")

terrestrial_pa <- test3

# Write the shape
st_write(terrestrial_pa, "data/spatial/protected_areas/Terrestrial_focal_areas.shp", overwrite=T, append=F)
```


```{r  ch2_35, warning=F, message=F, include=F }
terrestrial_pa <- st_read("data/spatial/protected_areas/Terrestrial_focal_areas.shp")
```

```{r  ch2_36}
plot(st_geometry(merged_shp), col="grey", border=F)
plot(st_geometry(terrestrial_pa), col="red", add=T)

```


### End nodes

End nodes represent where the connectivity paths will terminate.

#### High elevation protected areas

When we are measuring Ridge to Reef connectivity - we need to define a height threshold that represents a meaningful transition in elevation. What should this height be? We have initiated the analysis using a 1500m threshold, but why not use 1000m? Where we draw the line potentially influences the availability of high elevation areas which animals can move to. 

Below we provide graphical representation of different height thresholds across the area of interest:

```{r  ch2_37}
par(mfrow=c(1,1))
plot(st_geometry(merged_shp), col=rgb(0.95,0.95,0.95))
plot(dem_raster, las=1 ,breaks=c(500,1000,1500,2000, 2500, 3000), col=terrain.colors(5), bg="grey", add=T, main="")
legend("topright",c("0-500", "500-1000", "1000-1500", "1500-2000", "2000-2500", "3000+"), fill=c(rgb(0.95,0.95,0.95), terrain.colors(5)))

```

Examples of studies in central America examining elevation gradients:

Smith MA, Hallwachs W, Janzen DH (2014) Diversity and phylogenetic community structure of ants along a Costa Rican elevational gradient. Ecography (Cop) 37:720–731. https://doi.org/10.1111/j.1600-0587.2013.00631.x
  A study on ants - consdered "high elevation" to be between 1300 and 1600m and found high uniqueness in those high elevation sites.  
  
From: Neate-Clegg MHC, Jones SEI, Burdekin O, et al (2018) Elevational changes in the avian community of a Mesoamerican cloud forest park. Biotropica 50:805–815. https://doi.org/10.1111/btp.12596

"Over a 10-year period, we found general increases in avian species richness and diversity at mid-to-high elevations (>1200 m), but declines at low elevations. This suggests upslope shifts in the community with lowland biotic attrition (Colwell et al. 2008)"


```{r ch2_38}
sf_use_s2(FALSE)

high_raster <- dem_raster
# Make a binary raster
high_raster$SRTM90_V4.elevation.tif[high_raster$SRTM90_V4.elevation.tif<1500] <- NA
high_raster$SRTM90_V4.elevation.tif[high_raster$SRTM90_V4.elevation.tif>=1500] <- 1
#plot(high_raster)
# Convert to polygon

high_poly = st_as_sf(high_raster, as_points = FALSE, merge = TRUE)

#plot(st_geometry(high_poly), col="red")
names(high_poly) <- c("1500m_thresh", "geometry" )

# Clean up any issues
high_poly <- st_buffer(high_poly, 0)
focal_pa <- st_buffer(focal_pa, 0)

# To deal with the error
high_pa <- st_intersection(focal_pa, high_poly)

# plot(st_geometry(merged_shp))
# plot(st_geometry(high_poly), col="grey", border=F, add=T)
# plot(st_geometry(high_pa), add=T, col="green")

# Interactive plot to check
tmp <- st_union(high_pa)

leaflet() %>%
  addProviderTiles(providers$OpenTopoMap, group="Topo") %>%
  addProviderTiles(providers$Esri.WorldImagery, group="Satellite") %>%  # Add satellite data
  addPolygons(data=tmp, weight = 2, fillColor = "blue") %>%
  # Layers control
  addLayersControl(
    baseGroups = c("Satellite", "Topo"),
    options = layersControlOptions(collapsed = FALSE)
  )

st_write(high_pa, "data/spatial/area_of_interest/high_elevation_pas.shp", append=F)

```

We will use high elevation (>1500) protected areas as the "end" points for our connectivity analyses.

```{r ch2_39}
plot(st_geometry(high_pa))
plot(gfc_2020, las=1,nbreaks=11 ,breaks="equal", col=rev(terrain.colors(10)), bg="grey", main="", add=T)
plot(st_geometry(high_pa), add=T)

# This aggregate method is very slow. I need a different solution. 
#mean_cover <- aggregate(gfc_2020, high_pa[1,], FUN = mean) 
# Exactexactr
#mean_cover <- aggregate(gfc_2020, high_pa[2,], FUN = mean, na.rm = TRUE, exact = TRUE) 

```

### Existing corridors

#### Corridors according to CBM

Data obtained from the Central American Commission on Environment and Development (CCAD) - indirectly through PhD researcher Ruchi Patel <rdp20@psu.edu> - based on her paper: [Patel, R. (2021). Paper plans and possibility: A critical analysis of landscape conservation policy in the Mesoamerican Biological Corridor. Environmental Development, 37, 100600.](https://www.sciencedirect.com/science/article/pii/S2211464520301329)

The corridors span the following countries:

```{r ch2_40}
# Check this also overlaps with the MBC
CBM_shp<- st_read("data/spatial/CBM Regional/cbm_regional.shp", quiet = TRUE)
st_crs(CBM_shp) <- 4326
table(CBM_shp$PAIS)
corr_shp <- CBM_shp[CBM_shp$PAIS!="Mexico",]

plot(st_geometry(merged_shp))
#plot(st_geometry(CBM_shp), add=T, col="red")
plot(st_geometry(corr_shp), add=T, col=as.numeric(as.factor(corr_shp$PAIS)), border=F)
legend("topright",levels(as.factor(corr_shp$PAIS)), pch=15, col=1:length(levels(as.factor(corr_shp$PAIS))) )
```

### Other sources

#### Key Biodiversity Areas
Key Biodiversity Areas (KBAs) are sites of global significance for the conservation of biodiversity. Currently there are 15,524 KBAs acknowledged worldwide, and more are continue to be identified nationally using simple, globally standardised criteria and thresholds, based on biodiversity requiring safeguards at the site scale. There are 11 criteria organized into five categories, namely (1) threatened biodiversity, (2) geographically restricted biodiversity, (3) ecological integrity, (4) biological processes, and (5) irreplaceability. As the building blocks for designing the ecosystem, bottom-up approach and maintaining effective ecological networks, Key Biodiversity Areas are the starting point for landscape-level conservation planning. 

We will explore the intersection between our ridge-to-reef corridors and these designated areas. 

```{r, echo=F, warning=F}

```




Costa Rica: SINAC <http://www.sinac.go.cr/EN-US/correbiolo/Pages/default.aspx>


